### 2017-05-29 Supplement.tex
TableI and TableII in Manuscript.tex Modified. MCTracFolds.png, network.png, Stderr.pdf, meanE.pdf andSupplementary.tex created. Labels also modified on graphs

\section{Supplementary Material}

\begin{figure}
\includegraphics[width=14cm]{meanE}
\caption{ Correlation of average energy as calculated by ST and MS algorithms for our 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2. Figure shows that  ST and MS are consistent in their calculations. The errors associated with both algorithms are very small such that they are not visible on the scale of the average energy.}
\end{figure}